### 2020-11-29 Research Report - Boverhof/Tex report/main.tex

%Required packages
\usepackage[utf8]{inputenc}
\usepackage{apacite}
\usepackage{caption} 
\usepackage{setspace}

@@

A commonly employed measure of workload is the NASA-Task Load Index (or TLX) questionnaire, operationalizing workload in clusters of six different dimensions \cite{hart2006nasa}. A potential problem arousing with such measurements, is that they are conducted post-experiment, posing the risk of jeopardizing the quality of the data by introducing subjectivity. An example of such a danger resides in the observer bias, proving that actors participating in an experiment tend to over-exaggerate the treatment effect when having to report it themselves post-experiment \cite{mahtani2018catalogue}. An inherently more objective method for the assessment of MWL is to gather bio-signals during the experiment, and use these to classify the degree of perceived MWL. Examples of such bio-signals (hereafter modalities) include techniques such as electroencephalogram, eye tracking, galvanic skin response, functional near-infrared spectroscopy, etc. The advantage of an approach like this is that complementary information streams, each stemming from a different modality can be interpreted simultaneously \cite{ramachandram2017deep}. This yields an objective and rich multifaceted classification of a mental construct, such as MWL. This approach comes however at the cost of an increase in complexity. This complexity mainly resides in the proper designing of a complex framework that inputs the data from the utilized modalities, and outputs a single classification outcome.

The current research directly builds upon previous research conducted by Dolmans and colleagues, who proposed a framework for multi-modal deep-learning classification of MWL \cite{dolmans2020perceived}. The current research will explore the feasibility of a similar framework, but by adding to this a real-time component. This implies that classification is done whilst the experiment takes place, enabling the possibilities to alter the state of the experiment in real-time. Such an approach could enable a wide range of possibilities: consider for example a flight-simulation with the objective of training pilots. The possibility for such a simulation to play out dynamically could enhance the learning-experience, namely by catering the state of the simulation towards the the individual, based on a real-time classification of for example their MWL.

@@

In order to build both frameworks, three of the (arguably) most widely utilized modalities will be included. These modalities include the techniques of electroencephalogram (EEG), galvanic skin response (GSR) and photoplethysmography (PPG). It is important to stress again that the objective of the current research is not to gain insight into the most optimal model for analyzing data stemming from the previously delineated modalities. It is rather to build a (real-time) framework researchers can utilize, and to which they can flexibly add modalities conform their own research goals. Consequently, one of two design principles on which the architecture of the framework reclines is the concept of modularity. Modularity refers the extent to which different modalities can freely be added and/or removed, without the necessity to re-architect and rebuild the entire framework. The second adhered design principle is the principle generalizability, prescribing that the framework should not solely be utilizable in the context of MWL, but also for the measurement of other mental constructs.

@@

\subsubsection{First modality: Electroencephalogram (EEG)}
The first utilized modality is a technique  called EEG, which detects electrical activity in the brain using electrodes. EEG is a widely utilized method for classifying MWL during experiments. With a review of the complete literature on EEG classification with deep learning, Craik and colleagues found a total of 16 \% of all available papers to deal with MWL, lending credence to the widely employed phenomenon of EEG for investigating MWL \cite{craik2019deep}. Additionally, this review aimed to map the feasibility of using deep neural learning for a range of different EEG application separately. Summarizing the findings, Craik and colleagues report that studies mostly found deep belief networks and ConvNets to perform best when aiming to classify MWL, and advice one of these approaches consequently.\cite{craik2019deep}

Research by Schirrmeister and colleagues contrasted several differently designed convolutional neural networks (ConvNets) against the baseline method for EEG data (FBCSP) in order to decode imagined or executed tasks. The investigated networks included deep, shallow, deep-shallow hybrid and residual ConvNets. Both the deep and shallow ConvNets were found to reach at least similar, and in some regard better classification results, as compared with the FBCSP baseline model. Altogether, a deep ConvNet with four convolutional-max-pooling blocks was found to perform best, displaying an accuracy of 92.4 \% \cite{schirrmeister2017deep}. The appropriate design choices were found to be of importance in order to be able to reach this accuracy, including the employment of techniques such as batch normalization, dropout and the usage of the ELU activation function.

@@

A variant on the CovNet LSTM model was employed by Dolmans and colleagues, who aimed to classify MWL by means of amongst other modalities GSR (equally so as the current research) \cite{dolmans2020perceived}. The performance of this model was contrasted with a network consisting solely of fully connected dense layers. Conform with findings by Sun and colleagues, the hybrid model was found to perform best \cite{dolmans2020perceived}. The model architecture as utilized by Dolmans and colleagues deployed 2 convolutional layers, after each of which batch normalisation was performed. Subsequently, max pooling was performed, after which the network feeds into two LSTM layers, followed by a dense layer.

@@

\subsubsection{Framework Architecture}

\paragraph{Data fusion.}  
When architecting a multi-model framework, it is of importance to realize that the information stemming from the different modalities are required to be combined (i.e. fused) in order to ultimately result in a single classification. Fusion can be done at different time-points within the framework. Several fusion strategies as proposed by Ramachandram and colleagues will be considered \cite{ramachandram2017deep}.

Early (or data-level) fusion focuses on how to optimally combine data sources, before being fed into the classification model. Techniques that realize this include for example principle component analysis or factor analysis. Early fusing is usually very challenging, due to the fact that data stemming from different modalities differ substantially in terms of dimensionality and sampling rate. Another disadvantage of early fusing, is that usually the oversimplified assumption of conditional independence is made. This assumption is unrealistic in practice, for data stemming from different modalities are expected to be correlated \cite{ramachandram2017deep}.

@@

Lastly, intermediate fusion is the most widely employed fusion strategy for multi-modal deep-learning problems. Modalities are fused by simply adding a higher order layer, to which the individual deep-learning models separately defined for each modality feed into. This need not be a single layer, but could also be multiple layers, as long as each modality ultimately feeds into the highest order output layer \cite{ramachandram2017deep}.

\paragraph{Real-time component}
bla bla

@@

\subsection{Framework}
The architecture for the proposed framework is constructed by combining insights from the literature, as delineated in paragraph \ref{Relatedwork} with the formulated design principles in section \ref{Introduction} (i.e. the principles of molecularity and generalizability).

The utilized network for the EEG modality will be a ConvNet with four convolutional max-pooling blocks. The utilized network for the GSR modality will be a hybrid LSTM ConvNet hybrid.

>>>>

%Required packages
\usepackage[utf8]{inputenc}
\usepackage{apacite}
\usepackage{caption} 
\usepackage{setspace}
\usepackage{amsmath}

@@

A commonly employed measure of workload is the NASA-Task Load Index (or TLX) questionnaire, operationalizing workload in clusters of six different dimensions \cite{hart2006nasa}. Such measurements are usually conducted post experiment, which could introduce some bias. An example of such a bias is the observer bias, prescribing that actors participating in an experiment tend to over-exaggerate the treatment effect when having to report it themselves post-experiment \cite{mahtani2018catalogue}.

An alternative method for assessing MWL is by measuring bio-signals during the experiment, and use these to classify the degree of perceived MWL. Examples of such bio-signals (hereafter modalities) include techniques such as electroencephalogram, eye tracking, galvanic skin response, functional near-infrared spectroscopy, etc. When training such classification models, administration of a questionnaire such as the TLX is usually utilized for labeling. The advantage of an approach like this is that complementary information streams, each stemming from a different modality can be interpreted simultaneously \cite{ramachandram2017deep}. This yields an objective and rich multifaceted classification of a mental construct, such as MWL. Additionally, a separate model for each respondent individually can be fitted, catering towards the individual perception of MWL. This approach comes however at the cost of an increase in complexity. This complexity resides in the construction of a complex framework that inputs the data from the utilized modalities, and outputs a single classification outcome.

The current research directly builds upon previous research conducted by Dolmans and colleagues, who proposed a framework for multi-modal deep-learning classification of MWL \cite{dolmans2020perceived}. The current research will explore the feasibility of a similar framework, but by utilizing different modalities and data. Additionally, the current research will aim to incorporate a real-time component. This implies that classification is done whilst the experiment takes place, enabling the possibilities to alter the state of the experiment in real-time. Such an approach could enable a wide range of possibilities: consider for example a flight-simulation with the objective of training pilots. The possibility for such a simulation to play out dynamically could enhance the learning-experience, namely by catering the state of the simulation towards the the individual, based on a real-time classification of for example their MWL.

@@

In order to build both frameworks, three of the (arguably) most widely utilized modalities will be included. These modalities include the techniques of electroencephalogram (EEG), galvanic skin response (GSR) and photoplethysmography (PPG). It is important to stress that the objective of the current research is not to gain insight into the most optimal model for analyzing data stemming from the previously delineated modalities individually. It is rather to build a (real-time) framework researchers can utilize, and to which they can flexibly add modalities conform their own research goals. Consequently, one of two design principles on which the architecture of the framework reclines is the concept of modularity. Modularity refers the extent to which different modalities can freely be added and/or removed, without the necessity to re-architect and rebuild the entire framework. The second adhered design principle is the principle generalizability, prescribing that the framework should not solely be utilizable in the context of MWL, but also for the measurement of other mental constructs.

@@

\subsubsection{First modality: Electroencephalogram (EEG)}
The first utilized modality is a technique  called EEG, which detects electrical activity in the brain using electrodes. EEG is a widely utilized method for classifying MWL during experiments. With a review of the complete literature on EEG classification with deep learning, Craik and colleagues found a total of 16 \% of all available papers to deal with MWL, lending credence to the widely employed phenomenon of EEG for investigating MWL \cite{craik2019deep}. Additionally, this review aimed to map the feasibility of using deep neural learning for a range of different EEG application separately. Summarizing the findings, Craik and colleagues report that studies mostly found deep belief networks and ConvNets to perform best when aiming to classify MWL, and advice one of these approaches consequently \cite{craik2019deep}.

Research by \citeA{schirrmeister2017deep} contrasted several differently designed convolutional neural networks (ConvNets) against the baseline method for EEG data (FBCSP) in order to decode imagined or executed tasks. The investigated networks included deep, shallow, deep-shallow hybrid and residual ConvNets. Both the deep and shallow ConvNets were found to reach at least similar, and in some regard better classification results, as compared with the FBCSP baseline model. Altogether, a deep ConvNet with four convolutional-max-pooling blocks was found to perform best, displaying an accuracy of 92.4 \% \cite{schirrmeister2017deep}. The appropriate design choices were found to be of importance in order to be able to reach this accuracy, including the employment of techniques such as batch normalization, dropout and the usage of the ELU activation function.

A different approach is proposed by \citeA{tabar2016novel}, combining a ConvNet with a Stacked auto-encoder network (SAE). Within this network, the input layer feeds into a convolutional layer with the objective of learning the filters and network parameters. The output of this convolutional layer subsequently feeds into SAE part of the network, constituting an input layer, 6 hidden layers and an output layer. A classification accuracy of 90 \% was acquired with this model \cite{tabar2016novel} .

@@

A variant on the CovNet LSTM model was employed by \citeA{dolmans2020perceived}, who aimed to classify MWL by means of amongst other modalities GSR (equally so as the current research). The performance of this model was contrasted with a network consisting solely of fully connected dense layers. Conform with findings by Sun and colleagues, the hybrid model was found to perform best \cite{dolmans2020perceived}, displaying an accuracy of 82 \%. The model architecture as utilized by Dolmans and colleagues deployed 2 convolutional layers, after each of which batch normalization was performed. Subsequently, max pooling was performed, after which the network feeds into two LSTM layers, followed by a dense layer.

@@

\subsubsection{Fusion strategy}  
When architecting a multi-model framework, it is of importance to realize that the information stemming from the different modalities are required to be combined (i.e. fused) in order to ultimately result in a single classification. Fusion can be done at different time-points within the framework. Several fusion strategies as proposed by Ramachandram and colleagues will be considered \cite{ramachandram2017deep}.

Early (or data-level) fusion focuses on how to optimally combine data sources, before being fed into the classification model. Techniques that realize this include for example principle component analysis or factor analysis. Early fusing is usually challenging, especially in a multimodal scenario such as the current. This resides in the fact that data stemming from different modalities differ substantially in terms of dimensionality and sampling rate. Another disadvantage of early fusing, is that usually the oversimplified assumption of conditional independence is made. This assumption is unrealistic, for data stemming from different modalities are expected to be correlated in practice \cite{ramachandram2017deep}.

@@

Lastly, intermediate fusion is the most widely employed fusion strategy for multi-modal deep-learning problems. Modalities are fused by simply adding a higher order layer, to which the individual deep-learning models separately defined for each modality feed into. This need not be a single layer, but could also be multiple layers, as long as each modality ultimately feeds into the highest order output layer. The depth of the fusion (i.e. the amount of fusion layers) can be chosen conform the specific situation, posing intermediate fusion to be the most flexible, and therefore the most widely used fusion method \cite{ramachandram2017deep}.

\subsubsection{Real-time component}
bla bla

\subsubsection{Model optimization}
The technique of batch normalization was proposed by  \citeA{ioffe2015batch},  and is often applied in deep learning with the objective of enhancing the stability of a network. This is endeavored by including a batch normalization layer after each convolutional layer, re-centering and re-scaling the input feeding into this layer. If incorporating a batch normalization layer, it is recommended to do so before feeding into the activation function \cite{ioffe2015batch}. An increase in accuracy for EEG classification was attained by \citeA{dolmans2020perceived},  \citeA{schirrmeister2017deep} by specifying a batch normalization layer after each convolutional layer. Equally so, the best performing model for PPG data as proposed by \citeA{biswas2019cornet} included a batch normalization layer after each convolutional layer.

Pooling layers are often used in ConvNets, following a convolutional layer with the attempt to decrease the dimensionality. The objective of such layers are to merge similar features into one (for a more extensive elaboration see: \cite{lecun2015deep}). Considering the EEG ConvNets, both \citeA{schirrmeister2017deep} and \citeA{tabar2016novel} specified a max-pooling layer after each convolutional layer. The model proposed for GSR data by \citeA{sun2019hybrid} incorporate a max-pooling layer after each, but one of the convolutional layers. Lastly, the model as proposed for PPG data by \cite{biswas2019cornet}  specified a max pooling layer after each convolutional layer.

Hyper-parameter optimization (HPO) is a technique that can be used to optimize hyper-parameter including learning rate, dropout probability and momentum. The Optuna toolbox provides a method for creating a parameter search space, from which values for the hyper-parameters can be sampled \cite{akiba2019optuna}.

\subsection{Framework Architecture}
The architecture for the proposed framework is constructed by combining insights from the literature whilst keeping in mind the previously delineated design principles (i.e. the principles of modularity and generalizability).

The utilized network for the EEG modality is a ConvNet as proposed by \citeA{schirrmeister2017deep}. The network is designed to include four convolutional layers, each follow by a max pooling layer of stride $1* 3$. Ultimately, these blocks are followed up with a softmax layer in order end up with a classification. The Exponential Linear Unit (ELU) function is utilized as activation function, defined as:

\begin{equation}
R(x)= 
\begin{Bmatrix}
x & x \geq0 \\ 
\alpha (e^x -1) & x<0
\end{Bmatrix}
\bigskip
\end{equation}

The utilized network for the GSR modality is a LSTM ConvNet hybrid.

@@
